---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### Librarys
```{r message=FALSE}
library(data.table)
library(dplyr)
```

### Load data
```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

### Head
```{r}
head(data)
```


## What is mean total number of steps taken per day?

### Plot the histogram
```{r}
dataRmNA <- tbl_df(data) %>% filter(!is.na(steps)) %>% group_by(date) %>% summarise(Sum = sum(steps))

hist(dataRmNA$Sum,
     col=1, main="Histogram of total number of steps per day", xlab="Total number of steps in a day")
```

### Mean
```{r}
mean(dataRmNA$Sum)
```

### Median
```{r}
median(dataRmNA$Sum)
```
#### The mean is 10766.19 and the median is 10765

## What is the average daily activity pattern?

### Group by intenval and get mean from steps
```{r}
intervalRmNA <- tbl_df(data) %>% filter(!is.na(steps)) %>% group_by(interval) %>% summarise(Mean = mean(steps))
```
### Plot result
```{r}
plot(intervalRmNA$interval, intervalRmNA$Mean, type='l', col=1, 
     main="Average number of steps averaged over all days", xlab="Interval", 
     ylab="Average number of steps")
```
```{r}
maxInterval <- tbl_df(intervalRmNA) %>% filter(Mean == max(Mean))
```
#### The max mean steps is 206.1698 in interval 835

## Imputing missing values
```{r}
naRows <- tbl_df(data) %>% filter(is.na(steps))
```
### Get number of na rows
```{r}
nrow(naRows)
```
#### The Data set has 2304 na rows.
```{r}
fillNa <- data
for (i in 1:nrow(fillNa)){
  if (is.na(fillNa$steps[i])){
    interval_val <- fillNa$interval[i]
    row_id <- which(intervalRmNA$interval == interval_val)
    steps_val <- intervalRmNA$Mean[row_id]
    fillNa$steps[i] <- steps_val
  }
}
```
#### The na values are now equal to the mean of intervals.
```{r}
stepsFillNa <- tbl_df(fillNa) %>% group_by(date) %>% summarise(Sum = sum(steps))
hist(stepsFillNa$Sum, col=1, main="(Imputed) Histogram of total number of steps per day", xlab="Total number of steps in a day")
```

#### The median and mean are 10766.19.

## Are there differences in activity patterns between weekdays and weekends?


```{r}
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$week <- weekdays(data$date)
entreSemana <- tbl_df(data) %>% filter((week != "sábado" | week != "domingo") & !is.na(steps)) %>% group_by(interval) %>% summarise(Mean = mean(steps))
finSemana <- tbl_df(data) %>% filter((week == "sábado" | week == "domingo") & !is.na(steps)) %>% group_by(interval) %>% summarise(Mean = mean(steps))

par(mfrow=c(1,2))
plot(entreSemana$interval, entreSemana$Mean, type='l', col=1, 
     main="Weekday", xlab="Interval", 
     ylab="Mean steps")
plot(finSemana$interval, finSemana$Mean, type='l', col=1, 
     main="Weekend", xlab="Interval", 
     ylab="Mean steps")
```
